---
title: "Index"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("C:/Users/Work/Desktop/UpdatedMixOmicsPages/All Pages")

library(RWordPress)
library(XML)
library(RCurl)
library(knitr)
## functon to upload images
#opts_knit$set(upload.fun = imgur_upload, base.url = NULL) 
```

```{r}
# username = 'password'
options(WordpressURL = c('http://mixomics.org/xmlrpc.php'))
options(WordpressLogin = c( "maxbladen" = "" )) 

# note: you may have to play around with the following options 'http://mixomics.org/xmlrpc.php' (http, not https)
```

```{r}
# current page
knit2wp(input = '', 
        title = '',  
        shortcode = FALSE, publish = TRUE, action = c("newPage"))
```

GUIDELINES: the new files will be sent to drafts on wordpress

1 - go to the updated page in Pages/drafts, preview, then 'quick edit' then 'publish'

2 - trash the existing file in Pages/Published

3 - in Pages/published, then 'quick edit' then choose parent page (probably not super useful) 

  3.5 - Go to Appearance/Menus and adjust position there

4 - Go to Appearance/Menu, add the new page to menu (will be append at the bottom) then reorganise the tab then 'save menu'

5 - check on website


# LIST OF STUFF TO CHANGE:

<img src="" alt="newplot" style="max-width: 80%; display: block; margin: auto;"/>

<p style="text-align: center;"></p>

fig.align = "center",





**List of potential errors / things to check:**

- if SSL error, check the credentials, as it could be WordpressURL = 'http://mixomics.org/xmlrpc.php OR 'http://mixomics.org/xmlrpc.php' (http, not https) (for example Al is https, and KA is http)

- if Bad Request (HTTP 400) or time out, keep trying. The server is just not responding

- text in comment in the Rmd text (it is ok in the code)

- Headings very similar to the title. Remove any first heading in the document

- check references weblinks: maybe too long

- for CIM / networks, put out.width='30%' in the chunk
