---
#title: "Multiblock (s)PLS"
date: "`r Sys.Date()`"
output: html_document
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(dpi = 100, 
                      echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.show = TRUE, 
                      fig.keep = 'all', 
                      fig.align = "center",
                      out.width = "70%") 
```

# Quick Start

``` {r}
library(mixOmics) # import the mixOmics library
data(breast.TCGA) # extract the TCGA data

X1 <- breast.TCGA$data.train$mirna # use the mirna and mrna expression levels as 
X2 <- breast.TCGA$data.train$mrna  # the X datasets
X <- list(mirna = X1, mrna = X2)

Y <- breast.TCGA$data.train$protein # set the protein levels as the Y dataset
```

### Multiblock PLS
``` {r, eval = FALSE}
block.pls.result <- block.pls(X, Y, design = "full") # run the method

plotIndiv(block.pls.result) # plot the samples
plotVar(block.pls.result, legend = TRUE) # plot the variables
```

`?block.pls` can be run to determine all default arguments of this function:

* Number of components (`ncomp = 2`): The first two PLS components are calculated,
* Design matrix (`design = "full"`): The strength of all relationships between dataframes is maximised (= 1) - a "fully connected" design,
* PLS mode (`mode = regression`): A PLS regression mode is performed,
* Scaling of the data (`scale = TRUE`): Each block is standardised to zero means and unit variances.

### Multiblock sPLS
``` {r, eval = FALSE}
# set the number of features to use for the X datasets
list.keepX = list(mrna = rep(5, 2), mirna = rep(5,2)) 
 # set the number of features to use for the Y dataset
list.keepY = c(rep(10, 2))

block.spls.result <- block.spls(X, Y, design = "full", # run the method
                                keepX = list.keepX, keepY = list.keepY) 

# plot the contributions of each feature to each dimension
plotLoadings(block.pls.result, ncomp = 1) 
plotIndiv(block.pls.result) # plot the samples
plotVar(block.pls.result, legend = TRUE) # plot the variables
```

`?block.spls` can be run to determine all default arguments of this function:

* Same defaults as above for `block.pls`,
* Features to retain (`keepX`, `keepY`): If unspecified, these values will default to using all features of the original dataframes. 

# Multiblock (s)PLS

Note, "multiblock" will be abbreviated to "MB" within this page.

Prior to learning the functionality of the MB-(s)PLS methods, a solid understanding of standard [(s)PLS](http://mixomics.org/methods/spls/) is strongly advised. The MB variants are extensions of the (s)PLS methods for when more than two datasets are being assessed. This draws on the methodology of Generalised CCA [1]. Here, there are multiple predictor datasests (**X~1~**, ... **X~Q~**) and a response vector/matrix of continuous values (**y** / **Y**). As with the standard forms of these methods, MB-sPLS is the sparse variant of MB-PLS and uses feature selection when forming latent components. When dealing with high dimensional datasets, MB-sPLS would be the recommended method as the non sparse version suffers from a lack of interpretability in this contexts. 

MB-(s)PLS features the same four modes of operation as [(s)PLS](http://mixomics.org/methods/spls/), including "regression", "canonical", "invariant" and "classic". These each function the same way they do in (s)PLS. However, the `block.pls()` and `block.spls()` functions do not have the same tuning and performance assessment methods when compared to `pls()`, `spls()`, `block.plsda()` and `block.splsda()`.


## Important Parameters

### *design* 

For a breakdown of how to construct the `design` matrix, refer to the [N-Integration Methods page](http://mixomics.org/methods/n-integration-methods/).

### *indY* vs *Y*

There are two ways in which the response dataset can be specified. If it is included in the list of datasets passed in via the `X` parameter, then `indY` passes the index of the desired Y dataframe to the function. In the below examples, the protein expression data is set to be the response dataframe:

``` {r, eval = FALSE}
X1 <- breast.TCGA$data.train$mirna 
X2 <- breast.TCGA$data.train$mrna
X3 <- breast.TCGA$data.train$protein
X <- list(mirna = X1, mrna = X2, protein = X3)

block.pls.result <- block.pls(X, indY = 3)
```

The alternative method is to have the desired Y dataframe totally separate to all the X datasets:

``` {r, eval = FALSE}
X1 <- breast.TCGA$data.train$mirna 
X2 <- breast.TCGA$data.train$mrna
X <- list(mirna = X1, mrna = X2)

Y <- breast.TCGA$data.train$protein

block.pls.result <- block.pls(X, Y = Y)
```


# References

1. [Tenenhaus A and Tenenhaus M. Regularized generalized canonical correlation analysis. Psychometrika, 76(2):257–284, 2011.](https://link.springer.com/article/10.1007%2Fs11336-011-9206-8)


