---
title: "Bioindicators ammonia"
author: "Olivier Chapleur"
date: ''
output:
  word_document:
    toc: yes
    toc_depth: '4'
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 4
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '4'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage


# Objectives

Different samples were produced in experiment testing the influence of ammonia (studies 1 to 4). 
The general purpose of this study is to assess if bioindicators of inhibition by ammonia can be identified within studies 1 and 2 and used to predict inhibition in samples of other studies (studies 3 and 4). 

As 16S data was generated with different sequencing methodologies, it was agregated at the genus level. 

# Data import

```{r importdata,echo=FALSE,message=FALSE,warning=FALSE}
rm(list=ls(all=TRUE))
#packages used to treat the data
library(mixOmics) 
library(Hotelling)
library(knitr)

#data of the different studies is imported
#data has been filtered previously (all genera with a relative abundance always inferior to 3% (in all samples) are removed)
load("data_prediction_data.RData")
#data and metadata of studies 1 and 2 and studies 3 and 4 are grouped together.
couleur_inhib=c("cornflowerblue", "grey") # colors used for the graphs

```

# Data transformation 

```{r,echo=FALSE}
#raw data are transformed to relative abundances and are CLR transformed. 
 cat("\n")
TSS.divide = function(x){
 x/sum(x)
} #function to calculate the relative abundance

# function is applied to each row for studies 1 and 2
pc_abundance_studies_1_2 = (apply(t(abundance_studies_1_2), 1, TSS.divide))
#clr transformation of data from studies 1 and 2
clr_abundance_studies_1_2=clr(t(pc_abundance_studies_1_2+0.001))
#same process for studies 3 and 4
pc_abundance_studies_3_4 = (apply(t(abundance_studies_3_4), 1, TSS.divide))
clr_abundance_studies_3_4=clr(t(pc_abundance_studies_3_4+0.001))

```




\newpage

# General view of the data with PCA



```{r, echo=FALSE, fig.height=6, fig.width=8}
pca=pca((clr_abundance_studies_1_2), scale=FALSE,center = TRUE,ncomp=5, logratio='none')

a=plotIndiv(pca,  group=metadata_studies_1_2$inhib_inoc , ind.names = F,legend=F,pch = as.numeric(factor(metadata_studies_1_2$experiment))+15,pch.levels=(metadata_studies_1_2$experiment), title="PCA",legend.title = "Inhibitor", legend.title.pch = "Experiment", size.legend = rel(2.4),
size.legend.title = rel(2.5), col.per.group =  couleur_inhib)

```

\newpage

# Identification of biomarkers without taking into account the batch effect: sPLSDA and associated methods



Sparse PLSDA (sPLSDA) enables to select the variables discriminant of the groups of samples

## sPLSDA tuning (how many variables must be kept?)

An error rate is calculated according to the number of variables kept on each component. Number of variables corresponding to the best error rate is selected. 

```{r ,echo=FALSE, eval = TRUE, message=FALSE, warning=FALSE, fig.height=6,fig.width=8}

splsda_tune = tune.splsda(X = clr_abundance_studies_1_2,Y=metadata_studies_1_2$inhib_inoc, ncomp = 2,logratio = "none",
                          multilevel = NULL, 
                          
                          test.keepX = c(seq(5,50, 1)), 
                          validation = c('Mfold'), 
                          folds = 5, 
                          dist = 'max.dist', 
                          nrepeat = 10,progressBar = FALSE)

```

 
```{r ,echo=FALSE,eval=TRUE, warning=FALSE, fig.height=6,fig.width=8}

cat("The plot shows the average error rate with respect to the keepX values tested:")
plot(splsda_tune)
```

After tuning the parameters, sPLSDA can be computed. 

```{r,eval=TRUE,echo=FALSE}
#number of variables and components kept
choice.ncomp = length(splsda_tune$choice.keepX)
choice.keepX = splsda_tune$choice.keepX

# the sPLS-DA
splsda = splsda(X = clr_abundance_studies_1_2, 
                    Y = metadata_studies_1_2$inhib_inoc,
                    ncomp = choice.ncomp,
                    keepX = choice.keepX,
                    logratio= "none")
```

## sPLSDA visualisation

```{r , echo= FALSE,eval=TRUE,fig.height=6,fig.width=8}




b=plotIndiv(splsda,  group=metadata_studies_1_2$inhib_inoc , ind.names = F,legend=F,pch = as.numeric(factor(metadata_studies_1_2$experiment))+15,pch.levels=(metadata_studies_1_2$experiment), title="sPLSDA",  col.per.group =  couleur_inhib)


```




###  sPLS-DA performance (prediction error rate)


``` {r, include = TRUE,eval=TRUE, fig.height=6, fig.width=8}
splsda_perf = perf(splsda, validation = 'Mfold', folds = 5, 
                   progressBar = FALSE, nrepeat = 20)
cat("Perf with Mfold validation, nrepat=20, folds=5")

splsda_perf$error.rate

#plot(splsda_perf)
```


\newpage

# Identification of biomarkers while taking into account the batch effect: MINT sPLSDA

```{r,echo=FALSE, fig.height=6, fig.width=8}

#the combined data set X of studies 1 and 2
X = clr_abundance_studies_1_2
#dim(X) 

# the outcome vector Y:  
Y =metadata_studies_1_2$inhib_inoc
#length(Y) 
#summary(Y)

# the vector indicating each independent study
study = unlist(list(metadata_studies_1_2$experiment))
study=droplevels(study)
# number of samples per study:
#summary(study)

# experimental design
table(Y,study)
```

### Tuning: how many components and variable to keep in the model? 

The tuning uses leave one out cross validation.

```{r,echo=FALSE, fig.height=6, fig.width=8}
tune.mint = tune(X = X, Y = Y, study = study, ncomp = 6, test.keepX = seq(1, 50, 1), 
method = 'mint.splsda', dist = "max.dist", progressBar = FALSE)

#plot(tune.mint, col = color.jet(6))
# optimal number of components
tune.mint$choice.ncomp #tune.mint$choice.ncomp 

# optimal keepX
tune.mint$choice.keepX


```

### Computation of MINT sPLSDA and projection of the samples on components 1 and 2

```{r}


mint.splsda.res = mint.splsda(X = X, Y = Y, study = study, ncomp = 3,  
                              keepX = tune.mint$choice.keepX)

# cat("Variables kept on comp 1")
# selectVar(mint.splsda.res, comp = 1)
# 
# cat("Variables kept on comp 2")
# selectVar(mint.splsda.res, comp = 2)
```

```{r,echo=FALSE, fig.height=6, fig.width=8}


c=plotIndiv(mint.splsda.res,  group=metadata_studies_1_2$inhib_inoc , ind.names = F,legend=T,pch = as.numeric(factor(metadata_studies_1_2$experiment))+15,pch.levels=(metadata_studies_1_2$experiment), subtitle="MINT sPLSDA",legend.title = "Inhibitor", legend.title.pch = "Experiment", size.legend = rel(2.4),
size.legend.title = rel(2.5),  col.per.group =  couleur_inhib)


```



```{r, eval=FALSE,echo=FALSE, fig.height=8, fig.width=24}
library(ggpubr)
tiff(filename = "figure_3.tiff", width = 5000, height = 1800, res = 400,pointsize=12)

ggarrange(a$graph, b$graph,c$graph,
                    labels = c("A", "B", "C"),
                    ncol = 3, nrow = 1,common.legend = TRUE, legend = "right")
dev.off
```


\newpage







### Perfomance of the MINT sPLSDA model

The performance of the MINT sPLS-DA model can be assessed with the \texttt{perf} function. (Since the previous tuning of the MINT sPLS-DA model was conducted with the distance \texttt{max.dist}, the same distance is used to assess the performance of the final model.)

```{r,echo=FALSE, fig.height=6, fig.width=8}
perf.mint = perf(mint.splsda.res, progressBar = FALSE)
#plot(perf.mint, col = color.mixo(1:3))
perf.mint$global.error
perf.mint$global.error$BER
```






# Prediction of Lufan and Peng data, model on one component

```{r, echo=FALSE, eval=TRUE}
## model with one component only
tune.mint = tune(X = X, Y = Y, study = study, ncomp = 1, test.keepX = seq(5, 50, 1), 
method = 'mint.splsda', dist = "max.dist", progressBar = FALSE)



mint.splsda.res = mint.splsda(X = X, Y = Y, study = study, ncomp = 1,  
                              keepX = tune.mint$choice.keepX)


# we predict lufan
test.predict <- predict(mint.splsda.res, newdata = clr_abundance_studies_3_4, dist = "max.dist",
                        study.test = factor(metadata_studies_3_4$experiment))
Prediction <- test.predict$class$max.dist[, 1]

# the confusion table compares the real subtypes with the predicted subtypes

cat("Summary of the results of the prediction. ")
get.confusion_matrix(truth = metadata_studies_3_4$inhib_inoc,
                     predicted = Prediction)


#metadata=read.csv("metadata_all_data_all_exp.csv",row.names=1)


cat("Detailed results of the prediction. ")


```

\newpage


# here I try to extract the coordinate of initial and predicted samples to draw a figure with everything



```{r, echo=FALSE, eval=TRUE}
## here model with two components

tune.mint = tune(X = X, Y = Y, study = study, ncomp = 2, test.keepX = seq(1, 50, 1), 
method = 'mint.splsda', dist = "max.dist", progressBar = FALSE)



mint.splsda.res = mint.splsda(X = X, Y = Y, study = study, ncomp = 2,  
                              keepX = tune.mint$choice.keepX)


# we predict lufan
test.predict <- predict(mint.splsda.res, newdata = clr_abundance_studies_3_4, dist = "max.dist",
                        study.test = factor(metadata_studies_3_4$experiment))
Prediction <- test.predict$class$max.dist[, 1]



predit=data.frame(test.predict$variates) ##coordonn?es des points pr?dits
construit=data.frame(mint.splsda.res$variates$X) ## coordonn?es des points utilis?s pour la construiction 






```



```{r, eval=FALSE}

background2=background.predict(object=mint.splsda.res, comp.predicted = 1, dist='max.dist', xlim=c(-3,3),ylim = c(-3,3))

```






```{r, eval=TRUE}

ordered_factor=factor( factor(metadata_studies_3_4$prediction), levels(factor(metadata_studies_3_4$prediction))[c(4,3,1,2,8,7,5,6)])

# for(i in 1:length(background2))
#             {
#                 if(!is.null(background2[[i]]))
#                     background2[[i]]=data.frame(id=i,col=names(background2)[i],
#                                                background2[[i]])
#             }
# 
#             background2 = do.call(rbind,background2)

#tiff(filename = "figure_5.tiff", width = 3000, height = 1800, res = 400,pointsize=12)


  ggplot(data=predit, aes(x=dim1, y=dim2))+
 # geom_polygon(data = background2,aes(x=Var1, y=Var2,fill = col), inherit.aes = FALSE, show.legend =TRUE)+
    scale_fill_manual(values = c("lavender", "grey95"), name='Prediction')+
  geom_point(aes(color=ordered_factor, shape=ordered_factor), size=2.5)+
    theme(legend.title=element_blank()) + 
  scale_color_manual(name = "Study & inhibition", values=c("grey50", "green","cornflowerblue", "darkblue","grey50" , "cornflowerblue", "darkblue", "green" ), labels=levels(droplevels(ordered_factor)))  + 
  scale_shape_manual(name = "Study & inhibition", values=c(15,15,15, 15, 8,8,8,8),  labels=levels(droplevels(ordered_factor)))  +
  labs( y="Component 2", x = "Component 1",color="Inhibition")+
  theme_bw()+
theme(strip.text= element_text(size=16, face='bold'))+
  geom_text(label=metadata_studies_3_4$prediction_code_plot,hjust=-0.4,vjust=1.2, cex=3)
 
#dev.off()
```
