---
title: "Identifying bioindicators of ammonia and phenol inhibition"
author: "Olivier Chapleur"
date: ""
output:
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '4'
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage


# Objectives

Different samples were produced in 2 in-house independant experiments testing the influence of ammonia and phenol on anaerobic digestion (studies 1 and 2). Detailed description of the studies if available in 

*Poirier S. and Chapleur O. Influence of support media supplementation to reduce the inhibition of anaerobic digestion by phenol and ammonia: Effect on degradation performances and microbial dynamics. Data in Brief. 2018; 19: 1733-1754.*

and 

*Poirier S. and Chapleur O. Inhibition of anaerobic digestion by phenol and ammonia: Effect on degradation performances and microbial dynamics. Data in Brief. 2018: 2235-2239.*


The general purpose of this study is to assess if bioindicators of inhibition by ammonia or phenol can be identified. 

In this study we show that a strong experimental effect exists ("batch effect"). However, it can be removed with statistical methods to focus on the effect of inhibitors only. 


# Data used for the study

## Loading of the data

```{r importdata,echo=FALSE,message=FALSE,warning=FALSE}
rm(list=ls(all=TRUE))
#packages used to treat the data
library(mixOmics)
library(Hotelling)
library(knitr)

#data of studies 1 and 2 is imported
load("mint_phenol_ammonia.RData")
#data has been filtered previously (all OTUs with a relative abundance always inferior to 3% (in all samples) are removed)

couleur_inhib=c("cornflowerblue", "grey", "darkorange") # colors used for the graphs
```

## Transformation of the data

```{r,echo=FALSE}
#raw data are transformed to relative abundances and are CLR transformed. 
 cat("\n")
TSS.divide = function(x){
 x/sum(x)
} #function to calculate the relative abundance

# function is applied to each row for studies 1 and 2
pc_abundance_studies_1_2 = (apply(t(abundance), 1, TSS.divide))
#clr transformation of data from studies 1 and 2
clr_abundance_studies_1_2=clr(t(pc_abundance_studies_1_2+0.01))
 
metadata_studies_1_2=metadata

```


\newpage


# General view of the data with principal component analysis (PCA)


```{r, echo=FALSE, fig.height=6, fig.width=8}
## pca is computed
pca=pca((clr_abundance_studies_1_2), scale=FALSE,center = TRUE,ncomp=5, logratio='none') 
#obtained pca is plotted
plotIndiv(pca,  group=metadata_studies_1_2$inhib_inoc , 
          ind.names = F,legend=T,
          pch = as.numeric(factor(metadata_studies_1_2$experiment))+15,
          pch.levels=(metadata_studies_1_2$experiment), 
          title="PCA",legend.title = "Inhibitor", 
          legend.title.pch = "Experiment", size.legend = rel(2.4),
          size.legend.title = rel(2.5), col.per.group =  couleur_inhib)
```

An important study effect is visible. 

\newpage

# Identification of biomarkers with sparse partial least square discriminant analysis (sPLSDA)



Sparse PLSDA (sPLSDA) enables to select variables discriminant of groups of samples. It does not specifically take into account a "study" effect.

More information is available in *Lê Cao K.A., Boitard S., and Besse P. Sparse PLS discriminant analysis: Biologically relevant feature selection and graphical displays for multiclass problems. BMC Bioinformatics. 2011; 12.*

## sPLSDA tuning (how many variables must be kept?)

An error rate is calculated according to the number of variables kept on each component. Number of variables corresponding to the best error rate is selected.  

```{r ,echo=FALSE, eval = TRUE, message=FALSE, warning=FALSE, fig.height=6,fig.width=8}

splsda_tune = tune.splsda(X = clr_abundance_studies_1_2,Y=metadata_studies_1_2$inhib_inoc, ncomp = 2,logratio = "none",multilevel = NULL, test.keepX = c(seq(5,50, 1)), validation = c('Mfold'), folds = 5, dist = 'max.dist',nrepeat = 10,progressBar = FALSE)

cat("Tune splsda with Mfold validation, folds=5, nrepat=20, ie dataset is divided in 5 groups of samples, 4 groups are used to build the model, the last one is used as training set. The 5 groups are then permuted and process iterated. 20 repeats means that all the process is repeated 20 times, ie that 20 partitions of samples in groups of 5 are tested. ")
```

 
```{r ,echo=FALSE,eval=TRUE, warning=FALSE, fig.height=6,fig.width=8}

cat("The plot shows the average error rate with respect to the keepX values tested:")
plot(splsda_tune)
```

After tuning the parameters, sPLSDA can be computed. 

```{r,eval=TRUE,echo=FALSE}
#number of components and variables kept after tuning.
choice.ncomp = length(splsda_tune$choice.keepX)
choice.keepX = splsda_tune$choice.keepX

# the sPLS-DA
splsda = splsda(X = clr_abundance_studies_1_2, Y = metadata_studies_1_2$inhib_inoc,ncomp = choice.ncomp,keepX = choice.keepX,logratio= "none")
```

## sPLSDA visualisation

```{r , echo= FALSE,eval=TRUE,fig.height=6,fig.width=8}

plotIndiv(splsda,  group=metadata_studies_1_2$inhib_inoc , ind.names = F,legend=T,pch = as.numeric(factor(metadata_studies_1_2$experiment))+15,pch.levels=(metadata_studies_1_2$experiment), title="sPLSDA",  col.per.group =  couleur_inhib)

```

Even if the different samples are well grouped based on the type of inhibitor, a "study" effect is still visible. 

##  sPLS-DA performance (prediction error rate)


``` {r, include = TRUE,eval=TRUE, fig.height=6, fig.width=8}
splsda_perf = perf(splsda, validation = 'Mfold', folds = 5, 
                   progressBar = FALSE, nrepeat = 20)
cat("Perf with Mfold validation, nrepat=20, folds=5")

splsda_perf$error.rate

```


\newpage

# Identification of biomarkers while taking into account the batch effect: multivariate integrative method (MINT) sPLSDA


MINT focuses on combining and integrating independent studies measured on the same variables. More information is available in *Rohart F., Eslami A., Matigian N., Bougeard S., and Lê Cao K.A. MINT: A multivariate integrative method to identify reproducible molecular signatures across independent experiments and platforms. BMC Bioinformatics. 2017; 18.*


```{r,echo=FALSE, fig.height=6, fig.width=8}
#information concerning the origin of the sample (study name) will be used in MINT
study = unlist(list(metadata_studies_1_2$experiment))
study=droplevels(study)
```


## Tuning: how many components and variable to keep in the model? 



```{r,echo=FALSE, fig.height=6, fig.width=8}
tune.mint = tune(X = clr_abundance_studies_1_2, Y = metadata_studies_1_2$inhib_inoc, study = study, ncomp = 6, test.keepX = seq(1, 50, 1), 
method = 'mint.splsda', dist = "max.dist", progressBar = FALSE)

plot(tune.mint, col = color.jet(6))
# optimal number of components
tune.mint$choice.ncomp #tune.mint$choice.ncomp 

# optimal keepX
tune.mint$choice.keepX

```

After tuning the parameters, MINT sPLSDA can be computed. 

```{r,echo=FALSE, fig.height=6, fig.width=8}
## Computation of MINT sPLSDA and projection of the samples on components 1 and 2

mint.splsda.res = mint.splsda(X = clr_abundance_studies_1_2, Y = metadata_studies_1_2$inhib_inoc, study = study, ncomp = 3,  
                              keepX = tune.mint$choice.keepX)

# cat("Variables kept on comp 1")
# selectVar(mint.splsda.res, comp = 1)
# 
# cat("Variables kept on comp 2")
# selectVar(mint.splsda.res, comp = 2)
```

## MINT sPLSDA visualisation

```{r,echo=FALSE, fig.height=6, fig.width=8}
plotIndiv(mint.splsda.res,  group=metadata_studies_1_2$inhib_inoc , ind.names = F,legend=T,pch = as.numeric(factor(metadata_studies_1_2$experiment))+15,pch.levels=(metadata_studies_1_2$experiment), subtitle="MINT sPLSDA",legend.title = "Inhibitor", legend.title.pch = "Experiment", size.legend = rel(2.4),
size.legend.title = rel(2.5),  col.per.group =  couleur_inhib)
```



\newpage

## Heatmap of the selected variables

A Clustered Image Map including the final gene signature is plotted (Euclidian distance and ward linkage). 


```{r, eval=TRUE,echo=FALSE, fig.height=12, fig.width=9}

#a file heatmap_legend was created previously with the name of the OTUs to plot on the figure. 
heatmap_col_names=paste(colnames(clr_abundance_studies_1_2), heatmap_legend[colnames(clr_abundance_studies_1_2),1]) #creation of the labels to plot as a legend for the heatmap

color=c(colorRampPalette(c( 'darkblue',  'cornflowerblue'))(15),colorRampPalette(c('cornflowerblue','lemonchiffon', 'yellow'))(15), colorRampPalette(c(  'yellow', 'orange'))(40)) #color palette for the heatmap

cim_mint_splsda=cim( color=color,mat=mint.splsda.res, comp = c(1,2), 
    row.sideColors = couleur_inhib[as.numeric(metadata_studies_1_2$inhib_inoc)], row.names = FALSE,
    clust.method = c("ward", "ward"),xlab="OTUs", ylab="Samples", legend=list(cex=1.6), col.cex=0.9, margins=c(15,3), symkey=FALSE, col.names=heatmap_col_names)

# dev.off()
# # 
# tiff(filename = "figure_2bis.tiff", width = 3000, height = 3700, res = 400,pointsize=12)
# 
color=c(colorRampPalette(c( 'darkblue',  'cornflowerblue'))(15),colorRampPalette(c('cornflowerblue','lemonchiffon', 'yellow'))(15), colorRampPalette(c(  'yellow', 'orange', 'orangered', 'darkred'))(40))

cim_mint_splsda=cim( color=color,mat=mint.splsda.res, comp = c(1,2),
    row.sideColors = couleur_inhib[as.numeric(metadata_studies_1_2$inhib_inoc)], row.names = FALSE,
    clust.method = c("ward", "ward"),xlab="OTUs", ylab="Samples", legend=list(cex=1.6), col.cex=0.9, margins=c(15,3), symkey=FALSE, col.names=heatmap_col_names)

# dev.off()

```





### Perfomance of the MINT sPLSDA model

The performance of the MINT sPLS-DA model can be assessed with the \texttt{perf} function. (Since the previous tuning of the MINT sPLS-DA model was conducted with the distance \texttt{max.dist}, the same distance is used to assess the performance of the final model.)

```{r,echo=FALSE, fig.height=6, fig.width=8}
perf.mint = perf(mint.splsda.res, progressBar = FALSE)
#plot(perf.mint, col = color.mixo(1:3))
perf.mint$global.error
perf.mint$global.error$BER
```